### 2024-11-27 Manuscript/main_ecography_R1.tex
Update main_ecography_R1.tex

We applied all metrics to an empirical dataset of bird extinctions on six Hawaiian islands (Hawaii, Kauai, Lanai, Maui, Molokai, Oahu), an archipelago which has suffered large numbers of anthropogenic extinctions due to a range of extinction drivers \citep{Walther2022}.
A prehistoric species list (avifauna known to be present prior to human colonisation of the islands), historic species list (avifauna known to be present at $1500$ CE) and extant species list of the native community for each island were taken from \cite{matthews2023global} along with trait values for the extinct species (\textit{Asio flammeus} was removed from the historic and extant species lists as its colonisation status is uncertain).
We also added two marine species to the prehistoric species list; we compiled missing trait data for these two extinct species from \cite{sayol2021loss}.
We extracted traits for extant species from the AVONET database \citep{tobias2022avonet}.
We used $9$ morphological traits associated with dietary/foraging preference and dispersal ability \citep{pigot2020macroevolutionary,sheard2020ecological} to represent the functional diversity of the communities: mass; beak length (culmen); beak length (nares); beak width; beak depth; tarsus length; wing length; the length from the first secondary feather to the tip of the longest primary feather (Kipp's distance); and tail length.
Traits were log transformed and standardised to a mean of $0$ and standard deviation of $1$, prior to analyses.

@@

The empirical results are also intriguing in that the loss of Hawaiian bird's functional diversity due to extinction is not greater than expected by chance (at least for certain metrics; Figure \ref{Fig:empirical_results}) because one would expect that extinctions target specific areas of functional diversity space (i.e. birds with specific trait combinations).
This contrasts with studies that have shown that species with certain traits (e.g., larger species) are more likely to have gone extinct and thus that functional diversity loss is larger than through random extinctions (e.g. \citealt{sayol2021loss,Matthews2022}).
This illustrates the importance of positing the mechanism and process of interest as well as the pattern to capture it before analysing the results.
For example, for the divergence metrics using probability density, hypervolume or dendrogram methods, we recovered changes in trait space that are smaller than expected by chance when focusing on all extinctions and larger than expected by chance when considereing only pre-$1500$ extinctions.
% This could lead to the interpretation that the trait space was not changed under any specific mechanism before $1500$ (i.e. null mechanism) but that since then, specific mechanisms have been reducing specific areas of the trait space (e.g. filtering mechanism).
% However, when using a different metric (say the Rao's quadratic entropy based on distance matrices), the results would indicate that a reduction of specific areas of the trait space could have also occured before $1500$ leading to very different interpretation of the results.
%TM: can we make this interpretation? as this isn't the two time periods, i.e., there isn't a before and after 1500 comparison. It is all extinctions, and then just pre-1500 extinctions. 
%TM: And also see my previous comment - is not saying the opposite? That there was some mechanism pre 1500? But not when considering all extinctions?

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
 group & n & convex hull volume \\ 
  \hline
 current &  37 & 25.44 \\ 
 pre 1500 &  63 & 60.84 \\ 
   \hline
\end{tabular}
\caption{Trait space size measured as the convex hull volume for both groups.}
\label{Tab:results_convhull}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
 group & n & median displacement & 2.5\% & 25\% & 75\% & 97.5\% \\ 
  \hline
 current &  37 & 1.04 & 0.95 & 0.99 & 1.05 & 1.10 \\ 
 pre 1500 &  63 & 1.12 & 0.85 & 0.94 & 1.19 & 1.34 \\ 
   \hline
\end{tabular}
\caption{Trait space position distribution measured as the displacement metric: the percentages are the confidence intervals of the distribution.}
\label{Tab:displacement}
\end{table}

>>>>

We applied all metrics to an empirical dataset of bird extinctions on six Hawaiian islands (Hawaii, Kauai, Lanai, Maui, Molokai, Oahu), an archipelago which has suffered large numbers of anthropogenic extinctions due to a range of extinction drivers \citep{Walther2022}.
A prehistoric species list (avifauna known to be present prior to human colonisation of the islands), historic species list (avifauna known to be present at $1500$ CE) and extant species list of the native community for each island were taken from \cite{matthews2023global} along with trait values for the extinct species (\textit{Asio flammeus} was removed from the historic and extant species lists as its colonisation status is uncertain).
We also added two marine species to the prehistoric species list (\textit{Haliaeetus albicilla} and \textit{Pterodroma hypoleuca}); we compiled missing trait data for these two extinct species from \cite{sayol2021loss}.
We extracted traits for extant species from the AVONET database \citep{tobias2022avonet}.
We used $9$ morphological traits associated with dietary/foraging preference and dispersal ability \citep{pigot2020macroevolutionary,sheard2020ecological} to represent the functional diversity of the communities: mass; beak length (culmen); beak length (nares); beak width; beak depth; tarsus length; wing length; the length from the first secondary feather to the tip of the longest primary feather (Kipp's distance); and tail length.
Traits were log transformed and standardised to a mean of $0$ and standard deviation of $1$, prior to analyses.

@@

The empirical results are also intriguing in that the loss of Hawaiian bird's functional diversity due to extinction is not greater than expected by chance (at least for certain metrics; Figure \ref{Fig:empirical_results}) because one would expect that extinctions target specific areas of functional diversity space (i.e. birds with specific trait combinations).
This contrasts with studies that have shown that species with certain traits (e.g., larger species) are more likely to have gone extinct and thus that functional diversity loss is larger than through random extinctions (e.g. \citealt{sayol2021loss,Matthews2022}).
This illustrates the importance of positing the mechanism and process of interest as well as the pattern to capture it before analysing the results.
For example, for the divergence metrics using probability density, hypervolume or dendrogram methods, we recovered changes in trait space that are smaller than expected by chance when focusing on all extinctions and larger than expected by chance when considereing only pre-$1500$ extinctions.
% This could lead to the interpretation that the trait space was not changed under any specific mechanism before $1500$ (i.e. null mechanism) but that since then, specific mechanisms have been reducing specific areas of the trait space (e.g. filtering mechanism).
% However, when using a different metric (say the Rao's quadratic entropy based on distance matrices), the results would indicate that a reduction of specific areas of the trait space could have also occured before $1500$ leading to very different interpretation of the results.
%TM: can we make this interpretation? as this isn't the two time periods, i.e., there isn't a before and after 1500 comparison. It is all extinctions, and then just pre-1500 extinctions. 
%TM: And also see my previous comment - is not saying the opposite? That there was some mechanism pre 1500? But not when considering all extinctions?
%TODO@@@

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
 group & n & convex hull volume \\ 
  \hline
 current &  37 & 25.44 \\ 
 at 1500 &  63 & 60.84 \\ 
   \hline
\end{tabular}
\caption{Trait space size measured as the convex hull volume for both groups.}
\label{Tab:results_convhull}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
 group & n & median displacement & 2.5\% & 25\% & 75\% & 97.5\% \\ 
  \hline
 current &  37 & 1.04 & 0.95 & 0.99 & 1.05 & 1.10 \\ 
 at 1500 &  63 & 1.12 & 0.85 & 0.94 & 1.19 & 1.34 \\ 
   \hline
\end{tabular}
\caption{Trait space position distribution measured as the displacement metric: the percentages are the confidence intervals of the distribution.}
\label{Tab:displacement}
\end{table}
